### 2019-05-24 reviews/response.md

We thank the reviewers for their detailed and insightful feedback.

We have added four new implementations to Task Bench: Dask, bulk synchronous MPI,
MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
section, including preliminary results from experiments with GPUs on
the Piz Daint supercomputer.

Major changes in the revised paper:

  * New implementations: Dask, bulk synchronous MPI, MPI+OpenMP and MPI+CUDA
  * New sections:
      - Implementation: Sections 4.3 (Dask) and 4.5 (MPI+X)
      - Evaluation: Section 5.9 (GPU experiment)
  * New figures:
      - Listings 1 and 2 (code samples)
      - Figure 13 (GPU experiment)
  * Updated Figures 3 and 4 (see also lines 300-312)
  * Updated Section 7 with insights derived from Task Bench

@@

The memory bound kernel has been written to perform a varying number
of passes (or partial passes) over a fixed-sized buffer, in order to
keep working set size constant and avoid interference due to cache
effects, which in our opinion makes the results difficult to
interpret.

See the code samples in Listing 1 and 2.

Reviewer 2:

The conclusion has been updated with references to additional systems.

### 2019-05-25 reviews/response.md

We thank the reviewers for their detailed and insightful feedback.

We have added four new implementations to Task Bench: Dask, bulk synchronous MPI,
MPI+OpenMP and MPI+CUDA. Results are included in the evaluation
section, including preliminary results from experiments with GPUs on
the Piz Daint supercomputer.

Major changes in the revised paper:

  * New implementations: Dask, bulk synchronous MPI, MPI+OpenMP and MPI+CUDA
  * New sections:
      - Implementation: Sections 4.3 (Dask) and 4.5 (MPI+X)
      - Evaluation: Section 5.8 (GPU experiment)
  * New figures:
      - Listings 1 and 2 (code samples)
      - Figure 13 (GPU experiment)
  * Updated Figures 3 and 4 (see also lines 300-312)
  * Updated Section 7 with insights derived from Task Bench

### 2019-05-25 1_introduction.tex
Apply Alex's comments.

To address the first issue, we present Task Bench, a
\emph{parameterized} benchmark designed for exploring the performance
behaviors of parallel and distributed programming systems under a
variety of conditions. A critical property of Task Bench is that,
while able to express a wide variety of typical parallel programming
patterns, it is simple enough to be implemented with reasonable effort
to a level of quality sufficient for comparative
studies. Task Bench is provides a core API that is shared across the
various system implementations. New benchmarks created with Task Bench
immediately run on all systems, and new systems immediately run all
benchmarks. Thus, Task Bench is much more productive for benchmarking
a wide range of phenomena than the standard approach in which each
benchmark must be implemented separately for each system.

@@

To address the second issue and isolate the contribution of the programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG). Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
performance (e.g., FLOP/s or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
or strong scaling study, the behavior at the limit of scalability. For
weak scaling, METG(50\%) corresponds directly to the smallest problem size that can be weak-scaled with 50\% efficiency. For strong scaling, METG(50\%) can be used to compute the scale at which performance can be expected to dip below 50\% efficiency.

>>>>

To address the first issue, we present Task Bench, a
\emph{parameterized} benchmark designed for exploring the performance
behaviors of parallel and distributed programming systems under a
variety of conditions. A critical property of Task Bench is that,
while able to express a wide variety of typical parallel programming
patterns, it is simple enough to be implemented with reasonable effort
to a level of quality sufficient for comparative
studies. Task Bench provides a core API that is shared across the
various system implementations. New benchmarks created with Task Bench
immediately run on all systems, and new systems immediately run all
benchmarks. Thus, Task Bench is much more productive for benchmarking
a wide range of phenomena than the standard approach in which each
benchmark must be implemented separately for each system.

@@

To address the second issue and characterize the contribution of a programming model's runtime system to application performance, we introduce a new metric called \emph{minimum effective task granularity} (METG). Intuitively, METG(50\%) is the smallest executable task granularity that maintains at least 50\% efficiency, meaning that the application achieves at least 50\% of peak
performance (e.g., FLOP/s or memory bandwidth) on a given machine. METG captures the essence of what is important in a weak
or strong scaling study, the behavior at the limit of scalability. For
weak scaling, METG(50\%) corresponds directly to the smallest problem size that can be weak-scaled with 50\% efficiency. For strong scaling, METG(50\%) can be used to compute the scale at which performance can be expected to dip below 50\% efficiency.

### 2019-05-25 4_implementation.tex
Fix Dask citation.
Apply Alex's comments.

Dask is an implicitly parallel task-based programming system for large
scale data analytics with Python. Dask provides abstractions over
distributed NumPy arrays and Pandas dataframes, and also a lower-level
interface for launching tasks. We use the lower-level interface to
implement Task Bench.

@@

In our Task Bench implementation, we evaluate both PTG and DTD models,
as well as a third option based on DTD that has additional manual
optimizations applied to further reduce dynamic checks for what tasks
should be executed on a node, resulting in further improvements to
scalability.

>>>>

Dask~\cite{Dask15} is an implicitly parallel task-based programming system for large
scale data analytics with Python. Dask provides abstractions over
distributed NumPy arrays and Pandas dataframes, and also a lower-level
interface for launching tasks. We use the lower-level interface to
implement Task Bench.

@@

In our Task Bench implementation, we evaluate both PTG and DTD models,
as well as a third option based on DTD that has additional manual
optimizations applied to further reduce dynamic checks for what tasks
should be executed on a node, which further improves
scalability.

### 2019-05-25 5_evaluation.tex
Update radix numbers.
Apply Alex's comments.

The vast majority of systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOP/s compared
to systems which share all cores between the application and runtime. Some of the
slower systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this configuration ran for over 6 hours.

@@

%% From 1 to 3
%% dependencies, the ratio varies from $0.54\times$ to $7.3\times$ (median
%% $1.8\times$).

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $16\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a factor of $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible. For this reason, we recommend evaluating the performance of
programming systems under realistic workloads, rather than with
trivial dependencies.

@@

One of the basic questions when considering different programming
systems is: How much overhead does the system add? This question is tricky to answer directly because some systems introduce
overhead \emph{inline} (i.e., by running system internal processes on
the same cores as application tasks), while other systems introduce
overhead \emph{out-of-line} (i.e., by dedicating one or more cores
solely to runtime use). Some systems, like Charm++, Realm, and Regent,
even support both configurations.

To answer this question, we use the METG metric to determine the
smallest task granularity that can be executed at a given level of
efficiency as a proxy for overhead. Figure~\ref{fig:metg-compute}
shows how METG(50\%) varies with node count for a variety of
dependence patterns supported by Task Bench. METG(50\%) is calculated
separately at each node count, to isolate runtime system behavior from
changes in communication latencies and topology when using
progressively larger portions of the machine.

@@

METG is a useful metric in part because it summarizes the basic task
scheduling overhead of each programming system in a single number. This makes it
possible to evaluate METG at different node counts (shown in
Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
communication topology and latency. An interesting observation is that
the systems with the smallest METG on one node have roughly an order
of magnitude higher METG at 256 nodes. The obvious conclusion is that
increased communication latency requires significantly larger tasks to
achieve the same level of efficiency.

@@

The remaining differences are due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities. We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave analysis of persisitent load imbalance to future work.

@@

Figure~\ref{fig:cuda-efficiency} shows a comparison between the MPI
and MPI+CUDA implementations on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which compares favorably with the officially reported number
$4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
FLOP/s. Note that the CPU and GPU kernels perform different numbers of
operations as the GPU must be given more work to achieve peak
performance. The x-axis in Figure~\ref{fig:cuda-efficiency} has been
normalized so that the number of FLOPs is constant for a given problem
size.

Our MPI+CUDA implementation uses an offload model in which input and
output data is copied to and from the GPU on each time step. We tested
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
ranks per GPU to push work to the GPU in parallel and thus overlaps communication and computation. We see that the
\lstinline{w4} configuration achieves higher FLOP/s but
drops more rapidly at smaller problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. In either case, the
GPU requires more work to achieve high performance, and the overhead
of copying data to and from the GPU dominates at small task
granularities. Thus at smaller problem sizes the CPU achieves higher
performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
terms of MTEG (as the peak performance of the GPU and CPU are very
different), the conclusion here is similar to
Section~\ref{subsec:scalability}: The cost of communicating data and
tasks to and from the GPU imposes a higher task granularity to achieve
the same efficiency than running on the CPU, reducing the advantage of
MPI's very lightweight mechanisms.

>>>>

The vast majority of systems achieve or nearly achieve peak FLOP/s. Some
systems reserve a number of cores (usually 1 or 2) for internal
runtime usage; these systems take a minor hit in peak FLOP/s compared
to systems which share all cores between the application and runtime. Some of the
higher overhead systems struggle to achieve peak FLOP/s, though in most cases
the curves suggest that performance would continue to improve if we
were to run larger problem sizes. Unfortunately, the excessive
computational cost of running such tests makes this prohibitively
expensive. For example, the Spark job in
this configuration ran for over 6 hours.

@@

%% From 1 to 3
%% dependencies, the ratio varies from $1.0\times$ to $150\times$ (median
%% $2.3\times$).

The ratio in METG between 0 and 3 dependencies per task ranges from
$1.01\times$ to $250\times$ (median $2.9\times$). The difference is most pronounced in systems that
perform runtime work inline. For example, MPI achieves an METG of 390
ns with 0 dependencies, but this rises to 4.6 \textmu{}s with 3 dependencies,
a factor of $12\times$ increase. This is unsurprising as in the case
of 0 dependencies, no \lstinline[language=C++]{MPI_Isend} calls are
issued at all, so MPI has nothing to do aside from executing kernels
as quickly as possible. For this reason, we recommend evaluating the performance of
programming systems under realistic workloads, rather than with
trivial dependencies.

@@

One of the basic questions when considering different programming
systems is: How much overhead does the system add? This question is tricky to answer directly because some systems introduce
overhead \emph{inline} (i.e., by running system internal processes on
the same cores as application tasks), while other systems introduce
overhead \emph{out-of-line} (i.e., by dedicating one or more cores
solely to runtime use). Some systems, like Charm++, Realm, and Regent,
support both configurations.

To answer this question, we use the METG metric to determine the
smallest task granularity that can be executed at a given level of
efficiency as a proxy for overhead. Figure~\ref{fig:metg-compute}
shows how METG(50\%) varies with node count for a variety of
dependence patterns supported by Task Bench. METG(50\%) is calculated
separately at each node count, to distinguish runtime system behavior from
changes in communication latencies and topology when using
progressively larger portions of the machine.

@@

METG is a useful metric in part because it summarizes the basic task
scheduling overhead of each programming system in a single number. This makes it
possible to evaluate METG at different node counts (shown in
Figure~\ref{fig:metg-compute}) to see how it is impacted by changes in
communication topology and latency. An interesting observation is that
the systems with the smallest METG on one node have roughly an order
of magnitude higher METG at 256 nodes---increased communication latencies require significantly larger tasks to
achieve the same level of efficiency.

@@

The remaining differences are due primarily to
different scheduling behaviors. The execution of 4
simultaneous task graphs only partially mitigates the
load imbalance between tasks. Systems that provide an
additional on-node work stealing capability (such as Chapel with the distrib scheduler) see additional gains in
efficiency at large task granularities. However, the use of
work-stealing queues can also impact throughput at small task
granularities. For example, Chapel's default (non-work-stealing) scheduler outperforms Chapel distrib at very small task granularities. We do not consider Charm++ load balancers because the imbalance is \emph{non-persistent} (i.e., timestep $t$ is uncorrelated with timestep $t+1$). We leave analysis of persistent load imbalance to future work.

@@

Figure~\ref{fig:cuda-efficiency} shows a comparison between the MPI
and MPI+CUDA implementations on the Piz Daint supercomputer. The CUDA
compute kernel achieves a peak performance of $4.759 \times 10^{12}$
FLOP/s, which is very close to the officially reported number
$4.761 \times 10^{12}$. The CPU kernel achieves $5.726 \times 10^{11}$
FLOP/s. Note that the CPU and GPU kernels perform different numbers of
operations as the GPU must be given more work to achieve peak
performance. The x-axis in Figure~\ref{fig:cuda-efficiency} has been
normalized so that the number of FLOPs is constant for a given problem
size.

Our MPI+CUDA implementation uses an offload model in which input and
output data is copied to and from the GPU on each time step. We tested
two configurations: \lstinline{w1} uses 1 task per GPU, whereas
\lstinline{w4} overdecomposes by a factor of $4\times$, using 4 MPI
ranks per GPU to push work to the GPU in parallel and thus overlaps communication and computation. We see that the
\lstinline{w4} configuration achieves higher FLOP/s but
drops more rapidly at smaller problem sizes, due to the overhead of
scheduling $4\times$ as many CUDA kernel launches. In either case, the
GPU requires more work to achieve high performance, and the overhead
of copying data to and from the GPU dominates at small task
granularities, where the CPU achieves higher
performance. While Figure~\ref{fig:cuda-efficiency} is not couched in
terms of MTEG (as the peak performance of the GPU and CPU are very
different), the conclusion here is similar to
Section~\ref{subsec:scalability}: The cost of communicating data and
tasks to and from the GPU imposes a higher task granularity to achieve
the same efficiency than running on the CPU, reducing the advantage of
very lightweight mechanisms such as those in MPI.

### 2019-05-25 6_related.tex
Apply Alex's comments.

\textsc{coNCePTuaL}~\cite{Conceptual07} is a domain-specific language
for writing network performance tests. \textsc{coNCePTuaL} and Task
Bench both enable the easy creation of new benchmarks though
\textsc{coNCePTuaL} does so via scripting whereas Task Bench provides
a set of configurable parameters. \textsc{coNCePTuaL} also targets a
lower level of abstraction, optimized more for testing messaging
layers, whereas Task Bench is closer to application level and
therefore enables comparisons of a broader set of parallel and
distributed programming systems.

Limit studies of task scheduling throughput in various runtime systems
often make additional assumptions. A popular assumption is the use of
trivially parallel tasks~\cite{Canary16, Armstrong14}, which as shown
in Section~\ref{subsec:number-of-dependencies} underestimates (often
substantially) the cost of scheduling a task, and can also impact scalability.

>>>>

\textsc{coNCePTuaL}~\cite{Conceptual07} is a domain-specific language
for writing network performance tests. \textsc{coNCePTuaL} and Task
Bench both enable the easy creation of new benchmarks, though
\textsc{coNCePTuaL} does so via scripting whereas Task Bench provides
a set of configurable parameters. \textsc{coNCePTuaL} also targets a
lower level of abstraction, optimized more for testing messaging
layers, whereas Task Bench is closer to application level and
therefore enables comparisons of a broader set of parallel and
distributed programming systems.

Limit studies of task scheduling throughput in various runtime systems
often make additional assumptions. A popular assumption is the use of
trivially parallel tasks~\cite{Canary16, Armstrong14}, which as shown
in Section~\ref{subsec:number-of-dependencies} underestimates (often
substantially) the cost of scheduling a task and can also impact scalability.
